Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-08-01 manuscripts/reply/reply-main.tex
More tweaking of intro for reply ms.

Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
of vertebrate lineages distributed across the Philippines; they also studied
the behavior of the \msb approach using computer simulations.
They found the method is very sensitive to prior assumptions and often
supports shared divergences across taxa that diverged randomly over broad time
periods, to which \citet{Hickerson2013} responded.
\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
methodological point about the model selection performed in \msb:
\begin{itemize}
   \item The use of vague priors on the divergence-time parameters can lead to
       spurious support for models with few divergence events shared across
       taxa. Thus, the primary inference enabled by the approach is very
       sensitive to the priors on divergence times.
\end{itemize}
However, the two papers suggest alternative mechanisms by which the priors on
divergence times cause this behavior:
\begin{description}
    \item[Hypothesis 1: Numerical approximation error \citep{Hickerson2013}]
        Under broad uniform priors, the rejection algorithm implemented in \msb
        is unable to adequately sample the space of the models within
        reasonable computational time, which leads to bias toward models with
        fewer divergence-time parameters because they are better sampled.
    % \item \citet{Hickerson2013} suggest that under broad uniform priors, the
    %     rejection algorithm implemented in \msb is unable to adequately sample
    %     the space of the models within reasonable computational time, which
    %     leads to bias toward models with fewer divergence-time parameters
    %     because they are better sampled.
    \item[Hypothesis 2: Strongly weighted marginal likelihoods
        \citep{Oaks2012}]
        The uniform priors on divergence times simply lead to very small
        marginal likelihoods of models with many divergence-time parameters.
        These models are ``averaged'' over much greater parameter space with
        large prior weight and small likelihood
        \citep{Jeffreys1939,Lindley1957}.
    % \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
    %     priors on divergence times simply lead to very small marginal
    %     likelihoods of models with many divergence-time parameters. These
    %     models are ``averaged'' over much greater parameter space with large
    %     prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
\end{description}
In Hypothesis 1, the problem is numerical approximation error due to
insufficient computation;
% In the former case, the problem is numerical approximation error due to
% insufficient computation;
the exact posterior supports models with many divergence-time parameters
given data generated under temporally random divergences, but we are unable to
accurately approximate the posterior.
In Hypothesis 2, the problem is more fundamental, because the exact posterior
supports simultaneous divergences given data from randomly diverged taxa.
% In the latter case, numerical error is not the issue, because the exact
% posterior supports simultaneous divergences given data from randomly diverged
% taxa.
I.e., given the assumption that divergence times are broadly and uniformly
distributed, the exact posterior from Bayes' rule leads us to the wrong
conclusions about evolutonary history.
% I.e., this is the correct (albeit biogeographically misleading) answer given
% the prior assumption that divergence times are broadly and uniformly
% distributed.
Such posterior support for simultaneous divergence, even if ``correct'' from
the perspective of Bayesian model choice, does not provide the biogeographical
insights that a researcher who employs \msb seeks to gain.

While these phenomena are not mutually exclusive, it is important to
distinguish between them in order to determine how to improve our ability to
estimate shared divergence histories.
If Hypothesis 1 is correct, then the model is sound and we need to increase our
computational effort or improve our Monte Carlo integration procedures.
For example, Markov chain or sequential Monte Carlo algorithms might sample the
posterior more efficiently than the simple Monte Carlo rejection sampler
implemented in \msb.
Rather than alter the sampling algorithm, \citet{Hickerson2013} tried using
narrow, empirically informed uniform priors in the hope that with less
parameter space to sample, the rejection algorithm will produce better
estimates of the posterior.
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice and evaluate the approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their analyses, \citet{Hickerson2013} made an error by mixing different
units of time, which invalidates the results presented in their response (see
Supporting Information for details).
We correct this error, but still find their approach will often support
(1) clustered divergence models when divergences are random, and
(2) models that exclude from consideration the true values of the
parameters.
The behavior of their approach also suggests that insufficient computation
(Hypothesis 1) is not the primary cause of the issue.

>>>>

Recently, \citet{Oaks2012} used this model-choice framework to study 22 pairs
of vertebrate lineages distributed across the Philippines; they also studied
the behavior of the \msb approach using computer simulations.
They found the method is very sensitive to prior assumptions and often
supports shared divergences across taxa that diverged randomly over broad time
periods, to which \citet{Hickerson2013} responded.
\citet{Oaks2012} and \citet{Hickerson2013} agree on the fundamental
methodological point about the model selection performed in \msb:
\begin{itemize}
   \item The use of vague priors on the divergence-time parameters can lead to
       spurious support for models with few divergence events shared across
       taxa. Thus, the primary inference enabled by the approach is very
       sensitive to the priors on divergence times.
\end{itemize}
However, the two papers suggest alternative mechanisms by which the priors on
divergence times cause this behavior:
\begin{description}
    \item[Hypothesis 1: Numerical approximation error \citep{Hickerson2013}]
        Under broad uniform priors, the rejection algorithm implemented in \msb
        is unable to adequately sample the space of the models within
        reasonable computational time, which leads to bias toward models with
        fewer divergence-time parameters because they are better sampled.
    % \item \citet{Hickerson2013} suggest that under broad uniform priors, the
    %     rejection algorithm implemented in \msb is unable to adequately sample
    %     the space of the models within reasonable computational time, which
    %     leads to bias toward models with fewer divergence-time parameters
    %     because they are better sampled.
    \item[Hypothesis 2: Strongly weighted marginal likelihoods
        \citep{Oaks2012}]
        The uniform priors on divergence times simply lead to very small
        marginal likelihoods of models with many divergence-time parameters.
        These models are ``averaged'' over much greater parameter space with
        small likelihood, but large prior weight.
        \citep{Jeffreys1939,Lindley1957}.
    % \item \citet{Oaks2012} suggest the cause is more fundamental; the uniform
    %     priors on divergence times simply lead to very small marginal
    %     likelihoods of models with many divergence-time parameters. These
    %     models are ``averaged'' over much greater parameter space with large
    %     prior weight and small likelihood \citep{Jeffreys1939,Lindley1957}.
\end{description}
In Hypothesis 1, the problem is numerical approximation error due to
insufficient computation;
% In the former case, the problem is numerical approximation error due to
% insufficient computation;
the exact posterior supports models with many divergence-time parameters
given data generated under temporally random divergences, but we are unable to
accurately approximate the posterior.
In Hypothesis 2, the problem is more fundamental, because the exact posterior
supports simultaneous divergences given data from randomly diverged taxa.
% In the latter case, numerical error is not the issue, because the exact
% posterior supports simultaneous divergences given data from randomly diverged
% taxa.
I.e., given the assumption that divergence times are broadly and uniformly
distributed, the exact posterior from Bayes' rule leads us to the wrong
conclusions about evolutionary history.
% I.e., this is the correct (albeit biogeographically misleading) answer given
% the prior assumption that divergence times are broadly and uniformly
% distributed.
Such posterior support for simultaneous divergence, even if ``correct'' from
the perspective of Bayesian model choice, does not provide the biogeographical
insights that a researcher who employs \msb seeks to gain.

While these phenomena are not mutually exclusive, it is important to
distinguish between them in order to determine how to improve our ability to
estimate shared divergence histories.
If Hypothesis 1 is correct, then the model is sound and we need to increase our
computational effort or improve our Monte Carlo integration procedures.
For example, Markov chain or sequential Monte Carlo algorithms might sample the
posterior more efficiently than the simple Monte Carlo rejection sampler
implemented in \msb.
Rather than alter the sampling algorithm, \citet{Hickerson2013} tried using
narrow, empirically informed uniform priors in the hope that with less
parameter space to sample, the rejection algorithm will produce better
estimates of the posterior.
Here, we discuss theoretical considerations for using empirically informed
priors for Bayesian model choice and evaluate the approach of
\citet{Hickerson2013} as a potential solution to the biases of \msb revealed by
\citet{Oaks2012}.
In their analyses, \citet{Hickerson2013} made an error by mixing different
units of time, which invalidates the results presented in their response (see
Supporting Information for details).
We correct this error, but still find their approach will often support
(1) clustered divergence models when divergences are random, and
(2) models that exclude from consideration the true values of the
parameters.
% The behavior of their approach also suggests that insufficient computation
% (Hypothesis 1) is not the primary cause of the issue.
